---
title: "ch4 hw"
author: "Bruce Mallory"
date: "2/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 4.6
Suppose we collect data for a group of students in a statistics class with variables X1 = hours studied, X2 = undergrad GPA, and Y = receive an A. We fit a logistic regression and produce estimated coefficient, βˆ0 = −6, βˆ1 = 0.05, βˆ2 = 1.
(a) Estimate the probability that a student who studies for 40 h and has an undergrad GPA of 3.5 gets an A in the class.
(b) How many hours would the student in part (a) need to study to have a 50 % chance of getting an A in the class?
```{r cars}
beta.0 <- -6
beta.1 <- 0.05
beta.2 <- 1
hrs.a <- 40
gpa <- 3.5
regression <- beta.0 + beta.1*hrs.a + beta.2*gpa
prob.A <- exp(regression)/(1+exp(regression))
print(paste0("The probability of an A is ", 100*round(prob.A,2),"%"))

prob.B <- .50
log.odds <- (prob.B/(1-prob.B))
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
